---
title: "Simple Webservices with Plumber"
author: "Stefan Fleck"
date: "11/19/2020"
output: ioslides_presentation
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
library(here)
library(magrittr)
library(leaflet)
library(leaflet.extras)

knitr::opts_chunk$set(echo = FALSE, warning = FALSE)
```

## R Markdown



## Slide with Bullets

- Bullet 1
- Bullet 2
- Bullet 3


## The Problem:


```{r}
dd <- readRDS(here("data/brunn-am-gebirge_klagenfurt-am-woerthersee.rds"))

leaflet::leaflet(options = leaflet::leafletOptions(preferCanvas = TRUE))  %>%
  addFullscreenControl() %>% 
  leaflet::addProviderTiles(provider = "CartoDB.Positron") %>%
  leaflet::hideGroup("Route") %>%
  leaflet::addLayersControl(
    overlayGroups = c("Route", "Punkte")
  ) %>% 
  leaflet::addCircles(
    data = dd$points,
    color = "crimson",
    radius = 1,
    group = "Punkte"
  ) %>%
  leaflet::addPolylines(
    data = dd$sf, 
    group = "Route"
  )


```


## How do we 

* Manually call the script each time we receive a new router?
  - impractical
* Call an R script from the command line
  - 
* Use a web-service!


## What is a Web Service?

* A service that runs on a server and responds to requests
* Makes it possible 

```

client <-HTTP-> server

```


## What is a RESTful API

* API: Application Interface (as upposed to User Interface)
* REST: Representational state transfer (REST) is Standard for creating webservices.


## How is it useful?


```{R}
res <- HTTR::GET

```


ยง Further reading

* plumber: https://github.com/rstudio/plumber
* httr: https://github.com/r-lib/httr
* HTTP Methods explained: https://www.w3.org/Protocols/rfc2616/rfc2616-sec9.html
